### 2017-01-06 Dynamic Enumerator Allocation/Thesis/Version 2/bab2.tex

%-----------------------------------------------------------------------------%
\chapter{\babDua}
%-----------------------------------------------------------------------------%
\todo{tambahkan kata-kata pengantar bab 2 disini}