### 2017-04-28 Dynamic Enumerator Allocation/Thesis/Version 6/abstract.tex
Update thesis

%
% Halaman Abstract
%
% @author  Andreas Febrian
% @version 1.00
%

\singlespacing
\chapter*{ABSTRACT}
\medskip
\begin{center}
	{\large \textbf{CONTEXT-AWARE REAL-TIME LOCATION RECOMMENDATION SYSTEM}} \\
	\medskip
	By \\
	{\large \textbf{Aris Prawisudatama}} \\
	{\large \textbf{NIM: 23215131}} \\
	{\large \textbf{(Master of Electrical Engineering)}}
\end{center}
\medskip
\medskip
\medskip
\noindent Data collection is one of the main duties of a statistical agency in every country, including Badan Pusat Statistik (BPS). Practically, BPS use Census Block (BS) that is working area of every enumerator. Allocation of working area often done subjectively, causing imbalance completion time between enumerators. It will have an impact on the overall completion of the data collection. Despite the problem of enumerator allocation has similarities with MDVRP problem, MDVRP solver algorithm cannot directly solve this problem, because information of enumeration time is not available. \\

\noindent In this research, a system that can be used for generating recommendation of location is proposed. This system aims to provide a more equitable recommendation. Proposed system works gradually, by combining MDVRP solver algorithm with Publish/Subscribe mechanism. It will use context of each enumerator to make recommendation precisely.\\

\noindent Experiment was conducted by comparing proposed system and benchmark program, that is MDVRP algorithm without Publish/Subscribe mechanism. The result shows proposed system generate better recommendation in the most case.

\vspace*{0.2cm}

\noindent Keywords: recommendation; location routing; VRP; MDVRP; publish/subscribe \\

\onehalfspacing
\newpage